---
title: "Firearm Mortality_devon"
author: "Devon Morgan"
date: "11/13/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(usmap)
library(readxl)
library(patchwork)
library(viridis)
library(plotly)
library(rvest)
library(httr)
```

## Import and Tidy Firearm Mortality Data

Imported firearm mortality datasets from CDC Wonder database. 

```{r message = FALSE}

# Detailed Firearm Mortality Datatsets by Race, Age
firearm_mortality = read_csv("./data/cdc_firearm_mortality_data.csv", na = "Unreliable") %>% 
	janitor::clean_names() %>% 
	select(-ten_year_age_groups_code, -injury_mechanism_all_other_leading_causes_code, -race_code, death_cause= injury_mechanism_all_other_leading_causes) %>% 
	mutate(ten_year_age_groups = factor(ten_year_age_groups, levels = c("1-4 years", "5-14 years", "15-24 years", "25-34 years", "35-44 years", "45-54 years", "55-64 years", "65-74 years", "75-84 years", "85+ years")))

# Summary Firearm Mortality Dataset by Year
firearm_mortality_summary = read_excel("./data/cdc_firearm_all_ages.xlsx") %>% 
	janitor::clean_names() %>% 
	select(-year_code, -injury_mechanism_all_other_leading_causes_code, death_cause= injury_mechanism_all_other_leading_causes)

```

Added state abbreviations to Summary Firearm Mortality dataset: 

```{r }
# Created table with state name and state abbreviations crosswalk
st_crosswalk = tibble(state = state.name) %>%
   bind_cols(tibble(abb = state.abb)) %>% 
	 bind_rows(tibble(state = "District of Columbia", abb = "DC"))

# Joined abbreviation dataset with summary firearm mortality dataset
clean_firearm_mortality = left_join(firearm_mortality_summary, st_crosswalk, by = "state") %>% 
	rename(state_abb = abb)

```

## Import and Tidy Law Grade Dataset

```{r}
# website URL
url = "https://lawcenter.giffords.org/scorecard/#rankings"
gun_climate_url = read_html(url)

# extract table and clean data
gun_climate_data = 
	gun_climate_url %>% 
  html_nodes(css = "table") %>% 
  .[[1]] %>% 
  html_table(header = TRUE) %>% 
  as.tibble() %>% 
	janitor::clean_names() %>% 
	rename(
		law_strength = gun_law_strength_ranked, 
		grade_2017 = x2017grade, 
		death_rate_rank = gun_death_rate_ranked, 
		death_rate = gun_death_rate_per_100k) %>% 
	mutate(
    grade_2017 = factor(grade_2017, 
                      levels = c("A", "A-", "B+", "B", "C+", "C", "C-", "D", "D-", "F")), 
    grade_2017 = fct_collapse(grade_2017, 
    													A = c("A", "A-"), 
    													B = c("B+", "B"), 
    													C = c("C+", "C", "C-"), 
    													D = c("D", "D-")), 
    state_abb = state.abb) %>% 
		select(-state, -death_rate_rank, -death_rate)

```

## Merge Law Grade Dataset with Summary Firearm Mortality Dataset 

```{r}
merged_firearm_mortality = left_join(clean_firearm_mortality, gun_climate_data, by = "state_abb")
```


## Map of Crude Death Rates in 1999, 2007, 2012, 2016

```{r echo = FALSE}
merged_firearm_mortality %>% 
		filter(year == 1999) %>% 
		select(state_abb, crude_rate) %>% 
		rename(state = state_abb) %>% 
		plot_usmap(data = ., values = "crude_rate", lines = "red") + 
		scale_fill_continuous(
    low = "white", high = "red", name = "Crude Death Rate (1999)", label = scales::comma) +    
		theme(legend.position = "right")

merged_firearm_mortality %>%
		filter(year == 2003) %>% 
		select(state_abb, crude_rate) %>% 
		rename(state = state_abb) %>% 
		plot_usmap(data = ., values = "crude_rate", lines = "red") + 
		scale_fill_continuous(
    low = "white", high = "red", name = "Crude Death Rate (2003)", label = scales::comma) +    
		theme(legend.position = "none")

merged_firearm_mortality %>%
		filter(year == 2007) %>% 
		select(state_abb, crude_rate) %>% 
		rename(state = state_abb) %>% 
		plot_usmap(data = ., values = "crude_rate", lines = "red") + 
		scale_fill_continuous(
    low = "white", high = "red", name = "Crude Death Rate (2007)", label = scales::comma) +    
		theme(legend.position = "none")

merged_firearm_mortality %>%
		filter(year == 2012) %>% 
		select(state_abb, crude_rate) %>% 
		rename(state = state_abb) %>% 
		plot_usmap(data = ., values = "crude_rate", lines = "red") + 
		scale_fill_continuous(
    low = "white", high = "red", name = "Crude Death Rate (2012)", label = scales::comma) +    
		theme(legend.position = "none")

merged_firearm_mortality %>%
		filter(year == 2016) %>% 
		select(state_abb, crude_rate) %>% 
		rename(state = state_abb) %>% 
		plot_usmap(data = ., values = "crude_rate", lines = "red") + 
		scale_fill_continuous(
    low = "white", high = "red", name = "Crude Death Rate (2016)", label = scales::comma) +    
		theme(legend.position = "none")

```

## Spaghetti Plot of Crude Firearm Death Rate, by State, 1999-2016

```{r}

merged_firearm_mortality %>% 
  ggplot(aes(x = year, y = crude_rate, color = state_abb)) + 
  geom_line(alpha = 0.7, size = 0.5) +
  labs(
    title = "Crude Firearm Death Rate in Each State, 1999 to 2016",
    x = "Year",
    y = "Crude Death Rate",
    caption = "Data from the CDC Wonder Database"
  ) + 
  theme_bw() + 
	viridis::scale_color_viridis(discrete = TRUE,
                               name = "State")

```

## Crude Firearm Death Rate by Age Group and Race, 1999, 2016

```{r}
firearm_mortality %>% 
	filter(year == 1999) %>% 
	group_by(ten_year_age_groups, race) %>%
	summarize(n_deaths = sum(deaths),
						n_population = sum(population)) %>% 
	mutate(crude_rate = (n_deaths/n_population)*100000) %>% 
	ggplot(aes(fill= race, x = ten_year_age_groups, y = crude_rate, color = race)) +
	geom_bar(stat = "identity") +
	theme(axis.text.x = element_text(angle=60, hjust=1)) + 
	labs(
    title = "Crude Firearm Death Rate Across All States by Age Group and Race, 1999",
    x = "Age Group",
    y = "Crude Firearm Death Rate",
    caption = "Data from the CDC Wonder Dataset")

firearm_mortality %>% 
	filter(year == 2016) %>% 
	group_by(ten_year_age_groups, race) %>%
	summarize(n_deaths = sum(deaths),
						n_population = sum(population)) %>% 
	mutate(crude_rate = (n_deaths/n_population)*100000) %>% 
	ggplot(aes(fill= race, x = ten_year_age_groups, y = crude_rate, color = race)) +
	geom_bar(stat = "identity") +
	theme(axis.text.x = element_text(angle=60, hjust=1)) + 
	labs(
    title = "Crude Firearm Death Rate Across All States by Age Group and Race, 2016",
    x = "Age Group",
    y = "Crude Firearm Death Rate",
    caption = "Data from the CDC Wonder Dataset")



```
