---
title: "CDI"
author: 'Zanis Fang, UID: ZF2213'
date: "11/5/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

```

**Dataset Description:**

US Chronic Disease Indicators (CDI) dataset provides 124 chronic disease indicators provided by consensus. The source of the dataset is hosted on [data.gov](https://catalog.data.gov/dataset/u-s-chronic-disease-indicators-cdi). Major indicators from major categories were selected and joined with other variables for model building or visualization purposes. 

**Data cleaning:**
One major question from each chronic diseases were selected. Most indicators are from adults older than 18 years except for poverty (all age) and disability (older than 65 years.). The types of data values we use are the crude prevalence or mean, and we use overall data instead of stratified data. Data from year 2010 were removed because of huge amount of missing data.


```{r data_cleaning}
# list of questions
list_quest <- c(
"Poverty",
"Current smoking among adults aged >= 18 years",
"Prevalence of sufficient sleep among adults aged >= 18 years",
"Fair or poor self-rated health status among adults aged >= 18 years",
"Overweight or obesity among adults aged >= 18 years",
"No leisure-time physical activity among adults aged >= 18 years",
"Recent mentally unhealthy days among adults aged >= 18 years",
"Prevalence of diagnosed diabetes among adults aged >= 18 years",
"Heavy drinking among adults aged >= 18 years",
"Disability among adults aged >= 65 years")

# download the file if not downloaded yet
if (!file.exists("./data/U.S._Chronic_Disease_Indicators__CDI_.csv")) {
	download.file("https://chronicdata.cdc.gov/views/g4ie-h725/rows.csv?accessType=DOWNLOAD", "./data/U.S._Chronic_Disease_Indicators__CDI_.csv")
}
	
#
cdi_raw <- read_csv("./data/U.S._Chronic_Disease_Indicators__CDI_.csv")
cdi <- cdi_raw %>%
	# get geo locations and separate into latitude and longititude
	separate(GeoLocation, into = c("latitude", "longititude"), sep = ",") %>%
	mutate(latitude = str_replace(string = latitude, pattern = "\\(", replacement = ""),
				 longititude = str_replace(string = longititude, pattern = "\\)", replacement = ""),
				 latitude = as.numeric(latitude),
				 longititude = as.numeric(longititude)
	) %>% 
	# select the aimed questions
	filter(Question %in% list_quest) %>%
	# filtered out year 2010
	filter(LocationAbbr != "US",
				 YearStart != 2010,
				 Stratification1 == "Overall",
				 DataValueType %in% c("Crude Prevalence", "Mean")) %>%
	# get rid of unnecessary variables
	select(YearStart:DataValue, -DataSource, -Response, -DatavalueFootnote, -DataValueFootnoteSymbol, -YearEnd, -LocationDesc) %>%
	select(-Topic, -DataValueUnit, -DataValueType) %>% 
	# each column a question
	spread(key = Question, value = DataValue)

# label the column names
colnames(cdi) <- c("year", "state", "smoking", "disability_65", "self_rated_health", "drinking", "leisure_phys_act", "overweight", "poverty", "diabetes", "sleep", "mental_health")

# write the file for merging
if (!file.exists("./data/sel_cdi.csv")) {
	write_csv(cdi, "./data/sel_cdi.csv")
}

```

