---
title: "Socioeconomic status"
author: 'Zanis Fang, UID: ZF2213'
date: "11/14/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# loading necessary packages
library(tidyverse)

```


### Unemployment Data (State)

**Description:**

Monthly unemployment data of 51 states (including District of Columbia) from 1976 to 2018 were downloaded and extracted from [FRED economic research](https://fred.stlouisfed.org/search?st=unemployment+rate). Each dataset describe percentage of unemployment in each state, not seasonally adjusted.

**Cleaning:**

51 states were joined according to the date of unemployment rate. These datasets using four letter state abbreviations. Extract first two letters in order to compatible with other datasets. Year variable is extracted from the date variable. Yearly unemployment rates were averaged across months under the assumption that the population for each state holds constant. Tidy dataset was wrote into a "unempl.csv" for downstream analysis.




```{r}
# read all files
unempl_files <-
	tibble(state = list.files("./zanis_dataset/unemployment", full.names = TRUE)) %>%
	mutate(data = map(.x = state, .f = ~read_csv(.x))) %>%
	mutate(state = str_extract(state, "[A-Z]{4}"))

# join according to date
unempl <- unempl_files$data[[1]]
for (i in 2:nrow(unempl_files)) {
	unempl <- full_join(unempl, unempl_files$data[[i]], by = "DATE")
}

# get state abbreviations
unempl <- unempl %>% 
	gather(key = state, value = unemployment_rate, AKUR:WYUR) %>% 
	mutate(state = str_extract(state, "^[A-Z]{2}")) %>%
	mutate(year = lubridate::year(DATE)) %>% 
	group_by(state, year) %>% 
	# assuming the population hold constant acrossing the whole
	summarize(unemployment_rate = mean(unemployment_rate))

# write files if hasn't done yet
if (!file.exists("./zanis_dataset/unempl.csv")) {	
write_csv(x = unempl, path = "./zanis_dataset/unempl.csv")
}

```



