---
title: "Zanis Regression Report"
author: 'Zanis Fang, UID: ZF2213'
date: "12/3/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(PerformanceAnalytics)
library(modelr)
```

### Correlation Matrix

```{r correlation_table, message = FALSE, warning = FALSE}

data_for_reg <- read_csv("./zanis_dataset/merged_data.csv") %>%
	filter(year == 2016) %>% 
	spread(key = type_variable, value = Statistics) %>% 
	janitor::clean_names() %>% 
	filter(complete.cases(.)) %>% 
	mutate(law_strength_2016_only = 51 - law_strength_2016_only) %>% 
	select(-state, -year, -deaths)

chart.Correlation(data_for_reg)
```

*Comments:*

Strong correlations can be seen among several pairs of variables. To minimize multi-collinearity, model should be selected carefully.


### "Best" model at given number of variables

```{r generate_models, results = FALSE}
crude_rate_reg <- lm(data = data_for_reg, formula = crude_rate ~ .)
stepwise_lm <- step(crude_rate_reg, direction = "backward")

(criteria_reg <- leaps::regsubsets(crude_rate ~ ., data = data_for_reg) %>% summary())

```

```{r criteria_based}

criteria_df <- as_tibble(criteria_reg$outmat) %>%
	mutate("n_pred" = row_number()) %>%
	select(n_pred, everything()) %>% 
	t()

colnames(criteria_df) <- c(1:8)
criteria_df[-1, ] %>% knitr::kable()
	
	
```

*Comments:*




### R square, BIC, Cp, Regression Sum of Squares, Residual Sum of Squares

```{r model_selection}

tibble(
	n_pred = c(1:8),
	"Adjusted R-square" = criteria_reg$adjr2,
	"BIC" = criteria_reg$bic,
	"Cp" = criteria_reg$cp,
	"Regression Square Error" = criteria_reg$rsq,
	"Residual Sum Square" = criteria_reg$rss
	) %>%
	gather(key = "Statistics", value = "value", 2:6) %>%
	ggplot(aes(x = n_pred, y = value)) +
	  geom_point() +
	  geom_line() +
	  facet_grid(Statistics ~ ., scales = "free_y") +
	  labs(
	  	x = "Number of Predictors",
	  	y = "Values",
	  	title = "Model selection"
	  )
```

### Two selected models

**Model from stepwise:**

```{r stepwise}
summary(stepwise_lm) %>% broom::tidy() %>% knitr::kable()


```

**Criteria based model:**

```{r criteria}

criteria_lm <- lm(data = data_for_reg, formula = crude_rate ~ disability +
										law_strength_2016_only + smoking + unemployment_rate)
summary(criteria_lm) %>% broom::tidy() %>% knitr::kable()
```

### Model diagnostics

```{r model_diagnostics}
par(mfrow = c(2,5))
MASS::boxcox(criteria_lm)
plot(criteria_lm)
MASS::boxcox(stepwise_lm)
plot(stepwise_lm)
```

*Comments:*
No strong bias observed.

### Cross validation

```{r crossvalidation}


crossv_mc(data_for_reg, 100) %>% 
	mutate(train = map(train, as_tibble),
				 test = map(test, as_tibble)) %>%
	mutate(stpws_mdl = map(train, ~lm(crude_rate ~ drinking +
																			law_strength_2016_only +
																			leisure_physical_activities +
																			self_reported_health +
																			smoking +
																			unemployment_rate,
																		data = .x)),
				 triv_mdl = map(train, ~lm(crude_rate ~ 1,
																		data = .x)),
				 crt_mdl = map(train, ~lm(crude_rate ~ disability +
				 												 	law_strength_2016_only +
				 												 	smoking +
				 												 	unemployment_rate,
				 												 data = .x))) %>% 
	mutate(rmse_stpws = map2_dbl(stpws_mdl, test, ~rmse(model = .x, data = .y)),
				 rmse_triv = map2_dbl(triv_mdl, test, ~rmse(model = .x, data = .y)),
				 rmse_crt   = map2_dbl(crt_mdl, test, ~rmse(model = .x, data = .y))) %>% 
	select(starts_with("rmse")) %>%
	gather(key = model, value = rmse) %>% 
	ggplot(aes(x = model, y = rmse)) +
	  geom_violin()


```

